\begin{titlepage}

\begin{center}
\vspace{0.1in}
\large{\bf Curriculum Vitae}
\bigskip \bigskip
\end{center}

\begin{flushleft}
  {\bf Name:}{\hspace{3mm}}\fullname{}.\\
	{\bf Degree and date to be conferred:}{\hspace{3mm}}Ph.D. in Computer Science, May 2016. \\
	{\bf Collegiate institutions attended:}\\
	\begin{singlespace} 
	{\hspace{0.4in}}University of Maryland Baltimore County, Ph.D. Computer Science, 2015. \\
	{\hspace{0.4in}}\parbox[t]{5.5in}{Tulane University, M.S. Computer Science, 2000.} \\
        {\hspace{0.4in}}\parbox[t]{5.5in}{Tulane University, B.S. Computer Science, 1999.} \\
	\end{singlespace} 
	\vspace{8pt}
	{\bf Professional publications:}\\
	\begin{singlespace} 
{\hspace{0.4in}} \parbox[t]{5.5in}{B. Bauer, D. Scheidt, P. Rosendall, N. Rolander, \textbf{R. Holder}, E. Schmidt, and F. Ferrese.  ``Evaluating Test Methods for a Complex Connected System.''  American Institute of Aeronautics and Astronautics, March 2011.}\\
{\vspace{5pt}}

{\hspace{0.4in}} \parbox[t]{5.5in}{\textbf{R. Holder}.  ``Problem Space Analysis for Plan Library Generation and Algorithm Selection in
Real-time Systems.'' In Proceedings of The 23rd International FLAIRS Conference, May 2010, Daytona Beach, FL.  \textbf{Best Poster Award.}}\\
{\vspace{5pt}}

{\hspace{0.4in}} \parbox[t]{5.5in}{\textbf{R. Holder} and R.S. Cost.  ``Utilizing Resource Brokering Within Virtual Environments to Support Distributed Collaboration and Rapid Team Configuration.'' Proceedings of the 14th ICCRTS, June 2009, Washington, D.C.}\\
{\vspace{5pt}}

{\hspace{0.4in}} \parbox[t]{5.5in}{\textbf{R. Holder} and R.S. Cost.  ``Challenges of Resource Integration for C2 Collaborations within Virtual Environments.''  Proceedings of 2008 Proteus Futures Academic Workshop, Sept 2008, Carlisle, PA.}
\end{singlespace} 

\vspace{8pt}
	{\bf Professional positions held:}\\
	\begin{singlespace}
	{\hspace{0.4in}}\parbox[t]{5.5in}{Computer Scientist, Johns Hopkins University Applied Physics Laboratory (August 2000 -- December 2014).}\\
	{\hspace{0.4in}}\parbox[t]{5.5in}{Senior Software Engineer, ClearEdge IT Solutions, LLC (January 2015 -- Current).}\\	\end{singlespace}
\end{flushleft}
       
\end{titlepage}
\par\vfil


